---
title: "Model Fitting"
output: html_document
fig_width: 6 
fig_height: 4 
---

# 分縣市來做_密度_台東為例

## Referance

- http://idata8.com/rpackage/GWmodel/00Index.html **(GWmodel 的函數簡單說明) **
- https://www.datascienceblog.net/post/machine-learning/interpreting_generalized_linear_models/ **(glm residual )**
- https://stackoverflow.com/questions/2531489/understanding-glmresiduals-and-residglm
**Calling resid(model) will default to the deviance residuals, whereas model$resid will give you the working residuals. **
**resid(glm,type="working") = glm$residuals**
- **普瓦松回歸的 variance 會隨著y_hat 越大而越大(residual 也會越大)**
- https://online.stat.psu.edu/stat462/node/209/
- https://rpubs.com/benhorvath/glm_diagnostics

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=11, fig.height=4.5) 
```


## library the packages

```{r Warning = F}
library(GWmodel)      ### GW models
library(dplyr)
library(sp)           ## Data management
library(car)          ## vif
library(spdep)        ## Spatial autocorrelation
library(RColorBrewer) ## Visualization
library(classInt)     ## Class intervals
library(raster)       ## spatial data
library(grid)         # plot
library(gridExtra)    # Multiple plot
library(ggplot2)      # Multiple plot
library(gtable)
library(GGally)       # 相關係數圖矩陣（scatter plot matrix）
library(maptools)
library(MASS)
library(tmap)
library(gstat)        # Variogram
```

## Function 

### pdf_Plot

```{r pdf_Plot}
pdf_plot <- function(x, idx = c(rep(TRUE, length(x)))){
  g = ggplot()+
    geom_histogram(aes(x = x[idx], y = ..density..), 
                   fill = '#557C55', alpha = 0.8)+
    geom_density(aes(x = x[idx], y = ..density..), 
                 color = '#062C30', size = 1)+
    theme_bw()
  return(g)
}
```

### Point_Plot

```{r Point_Plot}
point_plot <- function(x,county, idx = c(rep(TRUE, length(x)))){
  idx.c = Variable_df$COUNTYNAME == county
  df = Variable_df[idx.c,]
  g = ggplot()+
    geom_point(aes(x = df$X[idx], y = df$Y[idx], color = x[idx] , size = abs(x[idx])))+
     scale_colour_gradient2(low = "#313552",mid = "#FBF8F1", high = "#632626")
  
  
  return(g)
}
```


## Load and tidy data

```{r Load data}
path = '..\\..\\Roaming-Dogs-Data\\'

# Variable_df <- read.csv(paste0(path, "@Taiwan_sampling_village_variable\\Taiwan_sampling_village_variable.shp" ), fileEncoding = 'utf-8')
Variable_shp<-shapefile(paste0(path, "@Taiwan_sampling_village_variable\\Taiwan_sampling_village_variable.shp" ),encoding = 'utf-8')
Variable_shp@data["School"] <- rowSums(Variable_shp@data[c("Ele","Junior" ,"Senior")])
Variable_shp@data["Student"] <- rowSums(Variable_shp@data[c("Ele_stu","Junior_stu" ,"Senior_stu")])
Variable_shp@data[is.na(Variable_shp@data)] <- 0

Variable_shp@data$Nr = Variable_shp@data$Nr %>% as.integer()
```

```{r Varaible name 01 original}
counties = "臺北市"
idx.county = Variable_shp@data$COUNTYNAME == counties
sum(idx.county)
col_X = 
  c( "Cluster",# 分群
     "Hospital","Clinic", "Temple", "School", "Train.stat", # 公共建設
     "Student", "Train.crow", # 人流
     "high_rat",   "mid_rat","low_rat", "M_F_RAT" ,"P_H_CNT","H_CNT", "P_DEN", "YOUN_DEP","OLD_DEP","AGING_IDX", # 人口統計(教育程度、人口密度...)
     "Income_mea","Income_sta", # 村里收入
     "Area_sqkm") 
col_y = c("Nr")
```

```{r 轉成數值}
for(i in c(col_X, col_y, "P_CNT", "X","Y")){
  
  Variable_shp@data[i] = sapply(Variable_shp@data[i], function(x) as.numeric(x))
}
```

## Clustering

分層：
- 原本的分層：依照人口密度、農業人口比例、專科以上人口比例及老年人口比例等變數，以群分析（Kmeans）作為都市化程度分層依據，依縣市為單位，將縣市內的的鄉鎮市區分為二至四層。

### Scale the Data
```{r scale the data}
# for (i in col_X[col_X != "Cluster"]){
# 
#     Variable_shp@data[i] <-scale(Variable_shp@data[i])
# 
# }
```

```{r Variable_df}
  
Variable_df = Variable_shp@data

```

```{r}
for (i in c(1:4)){
  
print(sum(Variable_df$Cluster[idx.county] == i))
  
}
```

```{r}
g = ggplot(data = Variable_df[Variable_df$COUNTYNAME == counties, ])+
  geom_boxplot(aes(x = as.factor(Cluster), y = Nr))
g
```


## Correlation
```{r correlation}
corr = cor(Variable_df[Variable_df$COUNTYNAME == "臺南市",c(col_y, col_X)])
idx = abs(corr[,col_y])>.15
col_X_02 = names(idx)[idx]; col_X_02 = col_X_02[col_X_02 != col_y]

corr[idx,idx]

```

```{r scatter plot}
ggpairs(Variable_df, columns =c(col_X_02,col_y))
```

## LASSO (Variables Selection，完全沒用！) 

<!-- - https://rpubs.com/chengjunhou/LASSO -->

<!-- ```{r} -->
<!-- library(glmnet) -->
<!-- X = Variable_df[col_X] %>% data.matrix() -->
<!-- y = Variable_df[col_y] %>% data.matrix() -->


<!-- fit = glmnet(X, y, family="poisson", nlambda=200, alpha=1, offset = log(Variable_df$Area_sqkm)) -->
<!-- print(fit) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- # coef(fit, s=c(fit$lambda[100],0.1)) -->
<!-- ``` -->


## Fitting the model

```{r X y offset}
col_X_03 = c("Cluster","School","Temple",  "P_H_CNT", "Income_mea", "Area_sqkm")
dt = Variable_df[idx.county,c(col_y, col_X_03)]
shp = Variable_shp[idx.county,c(col_y, col_X_03)]
print(dt %>% nrow())
f = Nr~.

ggpairs(Variable_df[idx.county,], columns =c(col_X_03,col_y))

residual_df = data.frame(VILLCODE = Variable_df$VILLCODE[idx.county],
                         COUNTYNAME = Variable_df$COUNTYNAME[idx.county],
                         TOWNNAME = Variable_df$TOWNNAME[idx.county],
                         VILLNAME = Variable_df$VILLNAME[idx.county],
                         X = Variable_df$X[idx.county],
                         Y = Variable_df$Y[idx.county],
                         Nr= Variable_df$Nr[idx.county],
                         Cluster = Variable_df$Cluster[idx.county],
                         P_H_CNT = Variable_df$P_H_CNT[idx.county],
                         Income_mea = Variable_df$Income_mea[idx.county],
                         Area_sqkm = Variable_df$Area_sqkm[idx.county])

```

### GLM_Poisson 
```{r glm.PO}
glm.PO <- glm(f,  family = poisson(link = "log"), data = dt)
summary(glm.PO)

fit.resp = predict(glm.PO, newdata = dt, type = "response") # 已用 exp 轉過的
fit.link = predict(glm.PO, newdata = dt, type = "link") # log(mu_i)
res.pear = residuals(glm.PO, type = "pearson")
res.work = residuals(glm.PO, type = "working")

residual_df$glm.PO.res.pear = res.pear
residual_df$glm.PO.fit = fit.resp

```
### GLM_NB
- https://bookdown.org/roback/bookdown-BeyondMLR/ch-poissonreg.html
- https://fukamilab.github.io/BIO202/04-C-zero-data.html

先只用各里寺廟、學校、診所密度與男女比、里平均收入建模

```{r glm.NB}
glm.NB<- glm.nb(f ,data = dt)

summary(glm.NB)

fit.resp = predict(glm.NB, newdata = dt, type = "response") # 已用 exp 轉過的
fit.link = predict(glm.NB, newdata = dt, type = "link") # log(mu_i)
res.pear = residuals(glm.NB, type = "pearson")
res.work = residuals(glm.NB, type = "working")

residual_df["glm.NB.res.pear"] = res.pear
residual_df["glm.NB.fit"] = fit.resp

```

### GWPR

- http://www-sre.wu.ac.at/ersa/ersaconfs/ersa13/ERSA2013_paper_01142.pdf
- 帶寬 (依照論文來定)

```{r}
DM<-gw.dist(dp.locat=data.matrix(((Variable_df[idx.county,c('X', "Y")]))))

# bw.gwr <- bw.ggwr(f,
#                  data = Variable_shp[c(col_X_03, col_y, "Area_sqkm")],
#                  family = "poisson",
#                  approach = "AICc",
#                  kernel = "gaussian",
#                  adaptive = TRUE,
#                  dMat = DM )
core.HR = mean(.3+.3+.33+.25)*10000; ext.HR = mean(7.7+5.7+5.7+5.6)*10000 
core.bw.gwr = (core.HR/pi)^.5; ext.bw.gwr = (ext.HR/pi)^.5; 
```

```{r}
gwr.PO <- ggwr.basic(f, 
                      data =shp,
                      family = "poisson",
                      bw = ext.bw.gwr, 
                      kernel = "gaussian",
                      adaptive = TRUE,
                      dMat = DM)
res.pear = (Variable_shp$Nr[idx.county]-gwr.PO$SDF$yhat)/(gwr.PO$SDF$yhat)^.5
residual_df["gwr.PO.res.pear"] = res.pear
residual_df["gwr.PO.fit"] = gwr.PO$SDF$yhat
summary(gwr.PO$SDF)
```



### GLMM_poisson with spatial (純粹看不懂) 

`glmmfields`:
- https://github.com/seananderson/glmmfields
- https://cran.r-project.org/web/packages/glmmfields/glmmfields.pdf
- http://idata8.com/rpackage/glmmfields/00Index.html
- https://www.mendeley.com/reference-manager/reader/ff1905f3-c0cf-38a2-a752-af1e7632d6eb/2c200ea5-4352-f33f-4851-2b712e71ce53
**Black swans in space: modeling spatiotemporal processes with extremes**
這篇論文在討論關於極端值的部分，但是似乎沒有使用

#### glmmfields
```{r glmmfields}
library(glmmfields)
# glm.PO.S <- glmmfields(f,
#   data = Variable_df[c(col_y, col_X_03, col_offset, "X", "Y")],
#   family = poisson(link = "log"),
#   lat = "Y", lon = "X",
#   nknots = 50,
#   iter = 500,
#   chains = 1,
#   prior_intercept = student_t(3, 0, 10),
#   prior_beta = student_t(3, 0, 3),
#   prior_sigma = half_t(3, 0, 3),
#   prior_gp_theta = half_t(3, 0, 10),
#   prior_gp_sigma = half_t(3, 0, 3),
#   seed = 123 # passed to rstan::sampling()
# )

# saveRDS(glm.PO.S, "glm_PO_S.rds")
glm.PO.S <- readRDS("glm_PO_S.rds")

```

```{r}
(glm.PO.S$model)
```
```{r}
fit.resp <-
predict(
glm.PO.S,
newdata = Variable_df[c(col_y, col_X_03, col_offset, "X", "Y")],
estimate_method = "median",
conf_level = 0.95,
interval =  "prediction",
type =  "response",
return_mcmc = FALSE,
iter = "all"
)

residual_df["glm.PO.S.res.resp"] = Variable_df$Nr-fit.resp$estimate
residual_df["glm.PO.S.res.pear"] = (Variable_df$Nr-fit.resp$estimate)/fit.resp$estimate^.5
residual_df["glm.PO.S.fit"] = fit.resp$estimate
```
```{r}
p1 = plot(glm.PO.S, type = "prediction", link = F) ; p2 = plot(glm.PO.S, type = "prediction", link = T)
ggpubr::ggarrange(p1, p2, ncol = 2)
```
```{r}
p1 = plot(glm.PO.S, type = "spatial-residual", link = T); p2 = plot(glm.PO.S, type = "spatial-residual", link = F)
ggpubr::ggarrange(p1, p2, ncol = 2)

```
```{r}
plot(glm.PO.S, type = "residual-vs-fitted", link = T);plot(glm.PO.S, type = "residual-vs-fitted", link = F)
```

#### spaMM

`spaMM`:
- https://kimura.univ-montp2.fr/~rousset/spaMM/spaMMintro.pdf 
- https://www.diva-portal.org/smash/get/diva2:685966/FULLTEXT02
**GLMM 與 GMRF 最大的差別就是後者的隨機效應並不獨立**
- https://www.r-bloggers.com/2019/09/spatial-regression-in-r-part-1-spamm-vs-glmmtmb/
**簡單介紹spaMM**

```{r spaMM poi}
library(spaMM)
set.seed(111)
n.train = as.integer(sum(idx.county)*0.8)
VILL.train = Variable_df$VILLCODE[idx.county] %>% sample(n.train)
VILL.test = Variable_df$VILLCODE[idx.county][!(Variable_df$VILLCODE[idx.county] %in% VILL.train)]

# Nr~Cluster+School+Temple+P_H_CNT+Income_mea+Area_sqkm+Matern(1|X+Y)
glmm_ComPo_C = fitme(Nr~P_H_CNT+Area_sqkm+Cauchy(1|X+Y),data= Variable_df[Variable_df$VILLCODE %in% VILL.train, c(col_y, col_X_03,  "X", "Y"),], 
      family = COMPoisson())
glmm_ComPo_M = fitme(Nr~P_H_CNT+Area_sqkm+Matern(1|X+Y),data= Variable_df[Variable_df$VILLCODE %in% VILL.train, c(col_y, col_X_03,  "X", "Y")], 
      family = COMPoisson())

```
```{r}
glmm_ComPo_C %>% summary() ; glmm_ComPo_M %>% summary()
```
```{r}

fit = predict(glmm_ComPo,  newdata = Variable_df[idx.county, c(col_X_03,  "X", "Y")],type = "response")

res.resp = residuals(glmm_ComPo,  type = "response")

residual_df$glmm_ComPo.fit = fit ;residual_df$glmm_ComPo.res.resp = res.resp
```
```{r}

qplot(x =  Variable_df[Variable_df$VILLCODE %in% VILL.train, "Nr"], y = residuals(glmm_ComPo_C,  type = "response")); 
qplot(x =  Variable_df[Variable_df$VILLCODE %in% VILL.train, "Nr"], y = residuals(glmm_ComPo_M,  type = "response"))
```
```{r}
predict(glmm_ComPo,  newdata = Variable_df[Variable_df$VILLCODE %in% VILL.test, c(col_X_03,  "X", "Y")],type = "response")
Variable_df[Variable_df$VILLCODE %in% VILL.test, "Nr"]
```

```{r spaMM NB}
library(spaMM)
col_X_03
# Nr~Cluster+School+Temple+P_H_CNT+Income_mea+Area_sqkm+Matern(1|X+Y)
glmm_NB = fitme(Nr~Cluster+P_H_CNT+Area_sqkm+Matern(1|X+Y),data= Variable_df[idx.county, c(col_y, col_X_03,  "X", "Y")], 
      family = negbin())
```
```{r}
glmm_NB %>% summary()

fit = predict(glmm_NB,  newdata = Variable_df[idx.county, c(col_X_03,  "X", "Y")],type = "response")

res.resp = residuals(glmm_NB,  type = "response")

qplot(x = c(fit), y = res.resp)

residual_df$glmm.NB.fit = fit ;residual_df$glmm.NB.res.resp = res.resp
```


### Other models
- file:///C:/Users/ChenWenRu/Downloads/imet131-j-chapitre-6.pdf

## Residual

用 pearson residual

Poisson Model:

$$r^d_i = \frac{{y_i} - \hat{\mu_i}}{\sqrt{\hat{\mu_i}}}$$
Negative Binomial Model:
- https://stats.stackexchange.com/questions/99406/why-are-pearsons-residuals-from-a-negative-binomial-regression-smaller-than-tho

$$r^d_i = ({{y_i} - \bar{\mu_i})}/{\sqrt{\bar{\mu_i}-\frac{\tilde{\mu}}{\theta}}}$$

### 觀察極端值

```{r glm.PO.res.pear}
idx.glm.PO = residual_df$glm.PO.res.pear>10
sum(idx.glm.PO)
df.02 = Variable_df[idx.county,]
outlier = df.02[idx.glm.PO , c("COUNTYNAME","TOWNNAME", "VILLNAME" ,col_y, col_X_03, "X", "Y")]
print(outlier)
```

### Plot

```{r Nr}
Nr = residual_df$Nr
p1 = pdf_plot(Nr,Nr<1000); p2 = point_plot(Nr,"臺南市",Nr<1000)
ggpubr::ggarrange(p1, p2, ncol = 2)

```

```{r glm.PO.res.pear}
res = residual_df$glm.PO.res.pear
p1 = pdf_plot(res,res<1000); p2 = point_plot(res,"臺南市",res<1000)
ggpubr::ggarrange(p1, p2, ncol = 2)
mean(res[res<100])
var(res[res<100])
```
```{r glm.PO.res.resp}
res = residual_df$Nr-residual_df$glm.PO.fit
p1 = pdf_plot(res); p2 = point_plot(res,"臺南市")
ggpubr::ggarrange(p1, p2, ncol = 2)
mean(res)
var(res)
```


```{r glm.NB.res.pear}
res = residual_df$glm.NB.res.pear
p1 = pdf_plot(res,res<500); p2 = point_plot(res,"臺南市",res<500)
ggpubr::ggarrange(p1, p2, ncol = 2)
```
```{r glm.NB.res.pear}
res = residual_df$Nr - residual_df$glm.NB.fit
p1 = pdf_plot(res); p2 = point_plot(res,"臺南市")
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r gwr.PO.res}
res = residual_df$gwr.PO.res.pear
p1 = pdf_plot(res); p2 = point_plot(res,"臺南市")
ggpubr::ggarrange(p1, p2, ncol = 2)
```


```{r glm.PO.S.res.pear}
res = residual_df$glm.PO.S.res.pear
p1 = pdf_plot(res,res<200); p2 = point_plot(res,res<200)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

### Variogram

```{r}
## Transform the dataframe into a SpatialPointDataFrame
## The objective is to obtain a spatial object
coordinates(residual_df)=~X+Y
class(residual_df)
```
```{r glm.PO.res.pear}
idx = residual_df$glm.PO.res.pear<1000
Vario_glm.PO = variogram(I(Nr-glm.PO.fit)~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_glm.PO)

```

```{r glm.NB.res.pear}
idx = residual_df$glm.NB.res.pear<1000
Vario_glm.NB=variogram(I(Nr-glm.PO.fit)~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_glm.NB)
```


```{r gwr.PO.res}
idx = residual_df$gwr.PO.res.pear>-900
Vario_gwr.PO=variogram(gwr.PO.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_gwr.PO)
```
```{r Nr}
idx = residual_df$gwr.PO.res.pear>-900
Vario_Nr =variogram(Nr~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_Nr)
```

### Check the original data

把殘差過大的地方拿出來看看

```{r}
villcode = residual_df$VILLCODE[residual_df$glm.NB.2.res>5]
dt = (Variable_df[Variable_df$VILLCODE %in% villcode ,c('VILLCODE','COUNTYNAME','TOWNNAME','VILLNAME',col_X_03, col_y)])

```