---
title: "Model Fitting"
output: html_document
fig_width: 6 
fig_height: 4 
---
## Referance

- http://idata8.com/rpackage/GWmodel/00Index.html **(GWmodel 的函數簡單說明) **
- https://www.datascienceblog.net/post/machine-learning/interpreting_generalized_linear_models/ **(glm residual )**
- https://stackoverflow.com/questions/2531489/understanding-glmresiduals-and-residglm
**Calling resid(model) will default to the deviance residuals, whereas model$resid will give you the working residuals. **
**resid(glm,type="working") = glm$residuals**
- **普瓦松回歸的 variance 會隨著y_hat 越大而越大(residual 也會越大)**
- https://online.stat.psu.edu/stat462/node/209/
- https://rpubs.com/benhorvath/glm_diagnostics

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=11, fig.height=4.5) 
```


## library the packages

```{r}
library(GWmodel)      ### GW models
library(dplyr)
library(sp)           ## Data management
library(car)          ## vif
library(spdep)        ## Spatial autocorrelation
library(RColorBrewer) ## Visualization
library(classInt)     ## Class intervals
library(raster)       ## spatial data
library(grid)         # plot
library(gridExtra)    # Multiple plot
library(ggplot2)      # Multiple plot
library(gtable)
library(GGally)       # 相關係數圖矩陣（scatter plot matrix）
library(maptools)
library(MASS)
library(tmap)
library(gstat)        # Variogram
```

## Function 

### pdf_Plot

```{r pdf_Plot}
pdf_plot <- function(x, idx = c(rep(TRUE, length(x)))){
  g = ggplot()+
    geom_histogram(aes(x = x[idx], y = ..density..), 
                   fill = '#557C55', alpha = 0.8)+
    geom_density(aes(x = x[idx], y = ..density..), 
                 color = '#062C30', size = 1)+
    theme_bw()
  return(g)
}
```

### Point_Plot

```{r Point_Plot}
point_plot <- function(x, idx = c(rep(TRUE, length(x)))){
  g = ggplot()+
    geom_point(aes(x = Variable_df$X[idx], y = Variable_df$Y[idx], color = x[idx] , alpha = .8))+
     scale_colour_gradient2(low = "#313552",mid = "#FBF8F1", high = "#632626")+
    theme_bw()
  
  
  return(g)
}
```


## Load and tidy data

分層：依照人口密度、農業人口比例、專科以上人口比例及老年人口比例等變數，以群分析（Kmeans）作為都市化程度分層依據分為二至四層。

```{r Load data}
path = '..\\..\\Roaming-Dogs-Data\\'

# Variable_df <- read.csv(paste0(path, "@Taiwan_sampling_village_variable\\Taiwan_sampling_village_variable.shp" ), fileEncoding = 'utf-8')
Variable_shp<-shapefile(paste0(path, "@Taiwan_sampling_village_variable\\Taiwan_sampling_village_variable.shp" ),encoding = 'utf-8')
Variable_shp@data["School"] <- rowSums(Variable_shp@data[c("Ele","Junior" ,"Senior")])
Variable_shp@data[is.na(Variable_shp@data)] <- 0

Variable_shp@data$Nr = Variable_shp@data$Nr %>% as.integer()
Variable_df = Variable_shp@data
```

```{r Varaible name 01 original}
col_X = 
  c( "Cluster", # 分群
     "Hospital","Clinic", "Temple", "Ele","Junior" ,"Senior", "Train.stat", # 公共建設
     "Ele_stu", "Junior_stu" ,"Senior_stu", "Train.crow", # 人流
     "high_rat",   "mid_rat","low_rat", "M_F_RAT" , "P_DEN", "YOUN_DEP","OLD_DEP","AGING_IDX", # 人口統計(教育程度、人口密度...)
     "Income_mea","Income_sta") # 村里收入
```

- 因為分群包含教育程度、老年人口比例與人口密度等資訊，因此將其相關變數去除。
- 因為里裡面的學生人數與扶幼率有極大正相關，因此使用扶幼率。
- 太多里沒有醫院，與之去除
```{r}
col_X = 
  c( "Cluster", # 分群
     "Clinic", "Temple", "School" ,"Train.stat", # 公共建設
     "Train.crow", # 人流
      "M_F_RAT" ,"YOUN_DEP", # 人口統計(教育程度、人口密度...)
     "Income_mea","Income_sta") # 村里收入
col_y = c('Nr')
```

```{r}
sum( Variable_df$Nr<250)
p1 = pdf_plot(Variable_df$Nr, idx = Variable_df$Nr<250) 
p2 = point_plot(Variable_df$Nr, idx = Variable_df$Nr<250)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r 轉成數值}
for(i in c(col_X, col_y, "P_CNT")){
  
  Variable_shp@data[i] = sapply(Variable_shp@data[i], function(x) as.numeric(x))
  Variable_df[i] = sapply(Variable_df[i], function(x) as.numeric(x))
}
```

### 轉成密度(公共建設分母為面積，遊蕩犬分母為面積)
```{r 轉成密度}
for(i in c( "Temple",  "School", "Train.stat",  "Clinic","Ele_stu", "Junior_stu" ,"Senior_stu", "Train.crow")){
  
  i_new = paste0(i, "_den")

  Variable_shp@data[i_new] = Variable_shp@data[i]/Variable_shp@data$Area_sqkm
  Variable_df[i_new] = Variable_df[i]/Variable_df$Area_sqkm
}
Variable_shp@data["Nr_den"] = Variable_shp@data["Nr"]/Variable_shp@data$Area_sqkm
Variable_df["Nr_den"] = Variable_df["Nr"]/Variable_df$Area_sqkm
```


```{r Varaible name 02}
col_X = 
  c( "Cluster","Temple_den",  "School_den", "Train.stat_den",  "Clinic_den", # 公共建設
     "Train.crow_den", 
     "M_F_RAT",  "YOUN_DEP", # 人口統計(教育程度、人口密度...)
     "Income_mea","Income_sta")
col_y= c('Nr')
col_y_den =c('Nr_den')
```

### Scale the Data
```{r scale the data}

# for (i in col_X[col_X != "Cluster"]){
# 
#     Variable_df[i] <- scale(Variable_df[i])
#     Variable_shp@data[i] <-scale(Variable_shp@data[i])
# 
# }
```

```{r Nr_den pdf}
idx_clu_1 = Variable_df$Cluster == 1
sum(Variable_df$Nr[!idx_clu_1]==0)/sum(!idx_clu_1); sum(Variable_df$Nr[idx_clu_1]==0)/sum(idx_clu_1); sum(Variable_df$Nr==0)/nrow(Variable_df)
pdf_plot(Variable_df$Nr[!idx_clu_1])+xlab("density of Nr(Cluster is 2,3,4 都市化較低)")
pdf_plot(Variable_df$Nr[idx_clu_1])+xlab("density of Nr (Cluster is 1 都市化較高)")
```

```{r}
g = ggplot(data = Variable_df)+
  geom_boxplot(aes(x = as.factor(Cluster), y = Nr))
g
```


## Correlation
```{r correlation Cluster == 1}
corr = cor(Variable_df[c(col_y, col_y_den,col_X)])
idx = abs(corr[,col_y_den])>.05 | abs(corr[,col_y])>.05
col_X_den = c("Temple_den",  "School_den", "Train.stat_den",  "Clinic_den")
corr[col_X_den,col_X_den]
col_X_02 = names(corr[idx,col_y_den]) %>% tail(-2)

corr[idx,idx]
```
```{r scatter plot}

ggpairs(Variable_df, columns =c(col_X_02,col_y_den)) 

```

```{r correlation Cluster != 1}
corr = cor(Variable_df[!idx_clu_1,c(col_y, col_y_den,col_X)])
idx = abs(corr[,col_y_den])>.1 | abs(corr[,col_y])>.1
col_X_den = c("Temple_den",  "School_den", "Train.stat_den",  "Clinic_den")
corr[col_X_den,col_X_den]
col_X_02 = names(corr[idx,col_y_den]) %>% tail(-1)

corr[idx,idx]
```


```{r scatter plot}

ggpairs(Variable_df, columns =c(col_X_02,col_y)) 

```

## LASSO (Variables Selection，完全沒用！) 

- https://rpubs.com/chengjunhou/LASSO

```{r}
library(glmnet)
X = Variable_df[col_X] %>% data.matrix()
y = Variable_df[col_y] %>% data.matrix()


fit = glmnet(X, y, family="poisson", nlambda=200, alpha=1, offset = log(Variable_df$Area_sqkm))
print(fit)
```
```{r}

coef(fit, s=c(fit$lambda[100],0.1))
```



## Fitting the model 
### GLM_Poisson 
```{r glm.PO}
col_X_03 = c("Temple_den","School_den", "Clinic_den","M_F_RAT","YOUN_DEP", "Income_mea")
f = Nr~Temple_den+School_den+Clinic_den+M_F_RAT+YOUN_DEP+Income_mea+offset(log(Area_sqkm))
col_offset = c("Area_sqkm")
glm.PO <- glm(f,  family = poisson(link = "log"), data = Variable_df[c(col_y, col_X_03, col_offset)])
summary(glm.PO)

fit.resp = predict(glm.PO, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "response") # 已用 exp 轉過的
fit.link = predict(glm.PO, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "link") # log(mu_i)
res.pear = residuals(glm.PO, type = "pearson")
res.work = residuals(glm.PO, type = "working")

residual_df = data.frame(VILLCODE = Variable_df$VILLCODE,
                         COUNTYNAME = Variable_df$COUNTYNAME,
                         TOWNNAME = Variable_df$TOWNNAME,
                         VILLNAME = Variable_df$VILLNAME,
                         X = Variable_df$X,
                         Y = Variable_df$Y,
                         Nr= Variable_df$Nr,
                         Cluster = Variable_df$Cluster,
                         glm.PO.res.pear =res.pear,
                         glm.PO.fit = fit.resp)
```
### GLM_NB
- https://bookdown.org/roback/bookdown-BeyondMLR/ch-poissonreg.html
- https://fukamilab.github.io/BIO202/04-C-zero-data.html

先只用各里寺廟、學校、診所密度與男女比、里平均收入建模

```{r glm.NB}
col_X_03 = c("Temple_den","School_den", "Clinic_den","M_F_RAT","YOUN_DEP", "Income_mea")
f = Nr~Temple_den+School_den+Clinic_den+M_F_RAT+YOUN_DEP+Income_mea+offset(log(Area_sqkm))
col_offset = c("Area_sqkm")
glm.NB <- glm.nb(f ,data = Variable_df[c(col_X_03, col_y, col_offset)])
# glm.NB.1 <- glm.nb(Nr ~.  + offset(off[idx_clu_1]),data = Variable_df[idx_clu_1,c(col_X_03, col_y)])
# glm.NB.2 <- glm.nb(Nr ~.  + offset(off[!idx_clu_1]),data = Variable_df[!idx_clu_1,c(col_X_03, col_y)])

summary(glm.NB)

fit.resp = predict(glm.NB, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "response") # 已用 exp 轉過的
fit.link = predict(glm.NB, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "link") # log(mu_i)
res.pear = residuals(glm.NB, type = "pearson")
res.work = residuals(glm.NB, type = "working")

residual_df["glm.NB.res.pear"] = res.pear
residual_df["glm.NB.fit"] = fit.resp

residual_df$glm.NB.res.pear = res.pear
```
### GLM_NB (reduce the X)
```{r glm.NB.02}
col_X_03 = c("Temple_den","Clinic_den","M_F_RAT","YOUN_DEP")
f = Nr~Temple_den+Clinic_den+M_F_RAT+YOUN_DEP+offset(log(Area_sqkm))
col_offset = c("Area_sqkm")
glm.NB.2 <- glm.nb(f ,data = Variable_df[c(col_X_03, col_y, col_offset)])
summary(glm.NB.2)

fit.resp = predict(glm.NB.2, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "response") # 已用 exp 轉過的
fit.link = predict(glm.NB.2, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "link") # log(mu_i)
res.pear = residuals(glm.NB.2, type = "pearson")
res.work = residuals(glm.NB.2, type = "working")


residual_df["glm.NB.2.res.pear"] = res.pear
residual_df["glm.NB.2.fit"] = fit.resp
```
### GWPR

- http://www-sre.wu.ac.at/ersa/ersaconfs/ersa13/ERSA2013_paper_01142.pdf

```{r}
col_X_03 = c("Temple","School", "Clinic","M_F_RAT","YOUN_DEP", "Income_sta")
off = log(Variable_df$Area_sqkm)
f = Nr~Temple+School+Clinic+M_F_RAT+YOUN_DEP+Income_sta+offset(log(Area_sqkm))

###
# 
# col_X_03 = c("Temple_den","Clinic_den","M_F_RAT","YOUN_DEP")
# f = Nr~Temple_den+Clinic_den+M_F_RAT+YOUN_DEP+offset(log(Area_sqkm))
# col_offset = c("Area_sqkm")
```

```{r}
DM<-gw.dist(dp.locat=data.matrix(((Variable_df[c('X', "Y")]))))

bw.gwr <- bw.ggwr(f,
                 data = Variable_shp[c(col_X_03, col_y, "Area_sqkm")],
                 family = "poisson",
                 approach = "AICc",
                 kernel = "gaussian",
                 adaptive = TRUE,
                 dMat = DM )
bw.gwr #20
```

```{r}
gwr.PO <- ggwr.basic(f, 
                      data =Variable_shp[c(col_X_03, col_y, "Area_sqkm")],
                      family = "poisson",
                      bw = 112, 
                      kernel = "gaussian",
                      adaptive = TRUE,
                      dMat = DM)
res.pear = (gwr.PO$SDF$yhat-gwr.PO$SDF$residual)/(gwr.PO$SDF$yhat)^.5
residual_df["gwr.PO.res.pear"] = res.pear
residual_df["gwr.PO.fit"] = gwr.PO$SDF$yhat
summary(gwr.PO$SDF)
```

### Mix GWPR (目前沒有 Poisson 的Function)
- https://iopscience.iop.org/article/10.1088/1742-6596/855/1/012007/pdf

```{r}
col_X_03 = c("Cluster","Temple_den","School_den", "Clinic_den","M_F_RAT","YOUN_DEP", "Income_sta")
off = log(Variable_df$Area_sqkm)
f = Nr~Cluster+Temple_den+School_den+Clinic_den+M_F_RAT+YOUN_DEP+Income_sta+offset(log(Area_sqkm))
```

```{r}
DM<-gw.dist(dp.locat=data.matrix(((Variable_df[c('X', "Y")]))))

bw.gwr <- bw.ggwr(f,
                 data = Variable_shp[c(col_X_03, col_y, "Area_sqkm")],
                 family = "poisson",
                 approach = "AICc",
                 kernel = "gaussian",
                 adaptive = TRUE,
                 dMat = DM )
bw.gwr #112
```

```{r}
mix.gwr.PO <- gwr.mixed(f, 
                      data =Variable_shp[c(col_X_03, col_y, "Area_sqkm")],
                      family = "poisson",
                      fixed.vars = c("Cluster","M_F_RAT","YOUN_DEP","Income_sta"),
                      bw = bw.gwr, 
                      kernel = "gaussian",
                      adaptive = TRUE,
                      dMat = DM)

summary(gwr.PO$SDF$yhat+ gwr.PO$SDF$residual)
```

```{r}
res.pear = (gwr.PO$SDF$yhat-gwr.PO$SDF$residual)/(gwr.PO$SDF$yhat)^.5

p1 = pdf_plot(res.pear,res.pear>-50 ); p2 = point_plot(res.pear,res.pear>-50)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r}
residual_df["gwr.PO.res"] = res.pear
residual_df["gwr.PO.fit"] = gwr.PO$SDF$yhat
```

### GLM_poisson with spatial (純粹看不懂) 

- https://cran.r-project.org/web/packages/glmmfields/glmmfields.pdf 
- https://kimura.univ-montp2.fr/~rousset/spaMM/spaMMintro.pdf 
- https://www.diva-portal.org/smash/get/diva2:685966/FULLTEXT02
**GLMM 與 GMRF 最大的差別就是後者的隨機效應並不獨立**
- https://www.r-bloggers.com/2019/09/spatial-regression-in-r-part-1-spamm-vs-glmmtmb/
**簡單介紹spaMM**

先只用各里寺廟、學校、診所密度與男女比、里平均收入建模
```{r glm.PO.S}
col_X_03 = c("Cluster","School_den", "Clinic_den","M_F_RAT", "Income_mea")
f = Nr~Cluster+School_den+Clinic_den+M_F_RAT+Income_mea+offset(log(Area_sqkm))
col_offset = c("Area_sqkm")

```
```{r}
library(glmmfields)
glm.PO.S <- glmmfields(f,
  data = Variable_df[c(col_y, col_X_03, col_offset, "X", "Y")],
  family = poisson(link = "log"),
  lat = "X", lon = "Y", nknots = 12, iter = 500, chains = 1,
  prior_intercept = student_t(3, 0, 10), 
  prior_beta = student_t(3, 0, 3),
  prior_sigma = half_t(3, 0, 3),
  prior_gp_theta = half_t(3, 0, 10),
  prior_gp_sigma = half_t(3, 0, 3),
  seed = 123 # passed to rstan::sampling()
)

summary(glm.PO.S$model@inits)
```
## Residual

用 pearson residual

Poisson Model:

$$r^d_i = \frac{\hat{y_i} - \hat{\mu_i}}{\sqrt{\hat{\mu_i}}}$$
Negative Binomial Model:
- https://stats.stackexchange.com/questions/99406/why-are-pearsons-residuals-from-a-negative-binomial-regression-smaller-than-tho

$$r^d_i = ({\hat{y_i} - \bar{\mu_i})}/{\sqrt{\bar{\mu_i}-\frac{\tilde{\mu}}{\theta}}}$$


### Plot

```{r glm.PO.res.pear}
res = residual_df$glm.PO.res.pear
p1 = pdf_plot(res); p2 = point_plot(res)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r glm.NB.res.pear}
res = residual_df$glm.NB.res.pear
p1 = pdf_plot(res); p2 = point_plot(res)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r glm.NB.2.res.pear}
res = residual_df$glm.NB.2.res.pear
p1 = pdf_plot(res); p2 = point_plot(res)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r gwr.PO.res}
res = residual_df$gwr.PO.res
p1 = pdf_plot(res); p2 = point_plot(res)
ggpubr::ggarrange(p1, p2, ncol = 2)
```


### Variogram

```{r}
## Transform the dataframe into a SpatialPointDataFrame
## The objective is to obtain a spatial object
coordinates(residual_df)=~X+Y
class(residual_df)
```
```{r glm.PO.res.pear}
idx = residual_df$glm.PO.res.pear<1000
Vario_glm.PO = variogram(glm.PO.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_RDT)
```

```{r glm.NB.res.pear}
idx = residual_df$glm.NB.res.pear<10
Vario_RDT=variogram(glm.NB.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_RDT)
```

```{r glm.NB.res.pear}
idx = residual_df$glm.NB.2.res.pear<10
Vario_RDT=variogram(glm.NB.2.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_RDT)
```

```{r gwr.PO.res}
idx = residual_df$gwr.PO.res.pear>-50
Vario_RDT=variogram(gwr.PO.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_RDT)
```

### Check the original data

把殘差過大的地方拿出來看看

```{r}
villcode = residual_df$VILLCODE[residual_df$glm.NB.2.res>5]
dt = (Variable_df[Variable_df$VILLCODE %in% villcode ,c('VILLCODE','COUNTYNAME','TOWNNAME','VILLNAME',col_X_03, col_y)])

```