---
title: "Model Fitting"
output: html_document
fig_width: 6 
fig_height: 4 
---
## Referance

- http://idata8.com/rpackage/GWmodel/00Index.html **(GWmodel 的函數簡單說明) **
- https://www.datascienceblog.net/post/machine-learning/interpreting_generalized_linear_models/ **(glm residual )**
- https://stackoverflow.com/questions/2531489/understanding-glmresiduals-and-residglm
**Calling resid(model) will default to the deviance residuals, whereas model$resid will give you the working residuals. **
**resid(glm,type="working") = glm$residuals**
- **普瓦松回歸的 variance 會隨著y_hat 越大而越大(residual 也會越大)**
- https://online.stat.psu.edu/stat462/node/209/
- https://rpubs.com/benhorvath/glm_diagnostics

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=11, fig.height=4.5) 
```


## library the packages

```{r}
library(GWmodel)      ### GW models
library(dplyr)
library(sp)           ## Data management
library(car)          ## vif
library(spdep)        ## Spatial autocorrelation
library(RColorBrewer) ## Visualization
library(classInt)     ## Class intervals
library(raster)       ## spatial data
library(grid)         # plot
library(gridExtra)    # Multiple plot
library(ggplot2)      # Multiple plot
library(gtable)
library(GGally)       # 相關係數圖矩陣（scatter plot matrix）
library(maptools)
library(MASS)
library(tmap)
library(gstat)        # Variogram
```

## Function 

### pdf_Plot

```{r pdf_Plot}
pdf_plot <- function(x, idx = c(rep(TRUE, length(x)))){
  g = ggplot()+
    geom_histogram(aes(x = x[idx], y = ..density..), 
                   fill = '#557C55', alpha = 0.8)+
    geom_density(aes(x = x[idx], y = ..density..), 
                 color = '#062C30', size = 1)+
    theme_bw()
  return(g)
}
```

### Point_Plot

```{r Point_Plot}
point_plot <- function(x, idx = c(rep(TRUE, length(x)))){
  g = ggplot()+
    geom_point(aes(x = Variable_df$X[idx], y = Variable_df$Y[idx], color = x[idx] , size = abs(x[idx])))+
     scale_colour_gradient2(low = "#313552",mid = "#FBF8F1", high = "#632626")
  
  
  return(g)
}
```


## Load and tidy data

```{r Load data}
path = '..\\..\\Roaming-Dogs-Data\\'

# Variable_df <- read.csv(paste0(path, "@Taiwan_sampling_village_variable\\Taiwan_sampling_village_variable.shp" ), fileEncoding = 'utf-8')
Variable_shp<-shapefile(paste0(path, "@Taiwan_sampling_village_variable\\Taiwan_sampling_village_variable.shp" ),encoding = 'utf-8')
Variable_shp@data["School"] <- rowSums(Variable_shp@data[c("Ele","Junior" ,"Senior")])
Variable_shp@data["Student"] <- rowSums(Variable_shp@data[c("Ele_stu","Junior_stu" ,"Senior_stu")])
Variable_shp@data[is.na(Variable_shp@data)] <- 0

Variable_shp@data$Nr = Variable_shp@data$Nr %>% as.integer()
```

```{r Varaible name 01 original}
col_X = 
  c( "Cluster", "Clu_new_2","Clu_new_3",  "Clu_new_4",  "Clu_new_5",# 分群
     "Hospital","Clinic", "Temple", "School", "Train.stat", # 公共建設
     "Student", "Train.crow", # 人流
     "high_rat",   "mid_rat","low_rat", "M_F_RAT" ,"P_H_CNT","H_CNT", "P_DEN", "YOUN_DEP","OLD_DEP","AGING_IDX", # 人口統計(教育程度、人口密度...)
     "Income_mea","Income_sta") # 村里收入
col_y = c("Nr")
```

```{r 轉成數值}
for(i in c(col_X, col_y, "P_CNT")){
  
  Variable_shp@data[i] = sapply(Variable_shp@data[i], function(x) as.numeric(x))
}
```

## Clustering

分層：
- 原本的分層：依照人口密度、農業人口比例、專科以上人口比例及老年人口比例等變數，以群分析（Kmeans）作為都市化程度分層依據，依縣市為單位，將縣市內的的鄉鎮市區分為二至四層。

- 我的分層：依照人口密度、、專科以上人口比例及老年指數等變數，以群分析（Kmeans）作為都市化程度分層依據，將全台灣的里分層。目前不知道要分幾層。

### 我的分層

```{r}
library(reshape2)
col_keams = c("P_DEN", "high_rat", "AGING_IDX")
col_clu = c("Cluster" ,"Clu_new_2","Clu_new_3",  "Clu_new_4",  "Clu_new_5")
df_cul = Variable_shp@data[c(col_keams, col_clu)]
for (i in col_keams){
    df_cul[i] <- scale(df_cul[i])
}
```

```{r}

colnames(df_cul)
df_cul_ori = melt(df_cul[c(col_keams ,col_clu[1])], id=col_clu[1])
df_cul_2 = melt(df_cul[c(col_keams ,col_clu[2])], id=col_clu[2])
df_cul_3 = melt(df_cul[c(col_keams ,col_clu[3])], id=col_clu[3])
df_cul_4 = melt(df_cul[c(col_keams ,col_clu[4])], id=col_clu[4])
df_cul_5 = melt(df_cul[c(col_keams ,col_clu[5])], id=col_clu[5])
colnames(df_cul_ori)

df_cul_ori = df_cul_ori %>% group_by(Cluster,variable) %>% summarise(mean_value = mean(value))
df_cul_2 = df_cul_2 %>% group_by(Clu_new_2,variable) %>% summarise(mean_value = mean(value))
df_cul_3 = df_cul_3 %>% group_by(Clu_new_3,variable) %>% summarise(mean_value = mean(value))
df_cul_4 = df_cul_4 %>% group_by(Clu_new_4,variable) %>% summarise(mean_value = mean(value))
df_cul_5 = df_cul_5 %>% group_by(Clu_new_5,variable) %>% summarise(mean_value = mean(value))
```

```{r}

ggplot(data = df_cul_ori, mapping = aes(x = variable, y = mean_value, linetype = as.factor(Cluster), colour = as.factor(Cluster), shape = as.factor(Cluster),group =  as.factor(Cluster)))+
  geom_line() + geom_point()

ggplot(data = df_cul_2, mapping = aes(x = variable, y = mean_value, linetype = as.factor(Clu_new_2), colour = as.factor(Clu_new_2), shape = as.factor(Clu_new_2),group =  as.factor(Clu_new_2)))+
  geom_line() + geom_point()

ggplot(data = df_cul_3, mapping = aes(x = variable, y = mean_value, linetype = as.factor(Clu_new_3), colour = as.factor(Clu_new_3), shape = as.factor(Clu_new_3),group =  as.factor(Clu_new_3)))+
  geom_line() + geom_point()

ggplot(data = df_cul_4, mapping = aes(x = variable, y = mean_value, linetype = as.factor(Clu_new_4), colour = as.factor(Clu_new_4), shape = as.factor(Clu_new_4),group =  as.factor(Clu_new_4)))+
  geom_line() + geom_point()

ggplot(data = df_cul_5, mapping = aes(x = variable, y = mean_value, linetype = as.factor(Clu_new_5), colour = as.factor(Clu_new_5), shape = as.factor(Clu_new_5),group =  as.factor(Clu_new_5)))+
  geom_line() + geom_point()
```

### 轉換 cul_new_3 Lable

原本的：

0 -> 都市化最低
1 -> 都市化中等
2 -> 都市化最低

要轉換成：

1 -> 都市化最高
2 -> 都市化中等
3 -> 都市化最低

```{r}
idx = Variable_shp@data$Clu_new_3

idx_new = c()

for (i in c(1:length(idx))){
  if(idx[i] == 0){
    idx_new[i] = 3
  }else if(idx[i] == 1){
    idx_new[i] = 2
  }else if(idx[i] == 2){
    idx_new[i] = 1
  }else{
    idx_new[i] = NA
  }
}

Variable_shp@data$Clu_new_3 = idx_new

```




### 轉成密度(公共建設分母為面積，遊蕩犬分母為面積)
```{r 轉成密度}
for(i in c( "Temple",  "School", "Train.stat",  "Clinic","Ele_stu", "Junior_stu" ,"Senior_stu", "Train.crow","H_CNT")){
  
  i_new = paste0(i, "_den")

  Variable_shp@data[i_new] = Variable_shp@data[i]/Variable_shp@data$Area_sqkm
}
Variable_shp@data["Nr_den"] = Variable_shp@data["Nr"]/Variable_shp@data$Area_sqkm
```

```{r}
Variable_df =Variable_shp@data 
```

- 因為分群包含教育程度、老年人口比例與人口密度等資訊，因此將其相關變數去除。
- 因為里裡面的學生人數與扶幼率有極大正相關，因此使用扶幼率。
- 太多里沒有醫院，與之去除
```{r}
col_X = 
  c( "Clu_new_3", # 分群
     "Clinic_den", "Temple_den", "School_den" ,"Train.stat_den", # 公共建設
     "Train.crow", # 人流
      "M_F_RAT" ,"YOUN_DEP","H_CNT_den", "P_H_CNT", # 人口統計(教育程度、人口密度...)
     "Income_mea","Income_sta") # 村里收入
col_y = c('Nr')
col_y_den =c('Nr_den')
```

### Scale the Data
```{r scale the data}

for (i in col_X[col_X != "Clu_new_3"]){

    Variable_df[i] <- scale(Variable_df[i])
    Variable_shp@data[i] <-scale(Variable_shp@data[i])

}
```


```{r}
g = ggplot(data = Variable_df)+
  geom_boxplot(aes(x = as.factor(Cluster), y = Nr))
g
g = ggplot(data = Variable_df)+
  geom_boxplot(aes(x = as.factor(Clu_new_3), y = Nr))
g

```


## Correlation
```{r correlation}
corr = cor(Variable_df[c(col_y, col_y_den,col_X)])
idx = abs(corr[,col_y_den])>.05 
col_X_den = c("Temple_den",  "School_den", "Train.stat_den",  "Clinic_den")
col_X_02 = names(corr[idx,col_y_den]) %>% tail(-2)

corr[idx,idx]
```
```{r scatter plot}
ggpairs(Variable_df, columns =c(col_X_02,col_y_den))
```

## LASSO (Variables Selection，完全沒用！) 

- https://rpubs.com/chengjunhou/LASSO

```{r}
# library(glmnet)
# X = Variable_df[col_X] %>% data.matrix()
# y = Variable_df[col_y] %>% data.matrix()
# 
# 
# fit = glmnet(X, y, family="poisson", nlambda=200, alpha=1, offset = log(Variable_df$Area_sqkm))
# print(fit)
```
```{r}
# coef(fit, s=c(fit$lambda[100],0.1))
```


## Fitting the model

```{r X y offset}
# 
# col_X_03 = c("Temple_den","School_den", "Clinic_den","M_F_RAT","YOUN_DEP", "Income_mea")
# f = Nr~Temple_den+School_den+Clinic_den+M_F_RAT+YOUN_DEP+Income_mea+offset(log(Area_sqkm))
# col_offset = c("Area_sqkm")

###
"Cluster"      

col_X_03 = c("Clu_new_3","Clinic_den", "Temple_den" )
f = Nr~Clu_new_3+Clinic_den+Temple_den+offset(log(Area_sqkm))
col_offset = c("Area_sqkm")

ggpairs(Variable_df, columns =c(col_X_03,col_y_den)) 
```

### GLM_Poisson 
```{r glm.PO}
glm.PO <- glm(f,  family = poisson(link = "log"), data = Variable_df[c(col_y, col_X_03, col_offset)])
summary(glm.PO)

fit.resp = predict(glm.PO, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "response") # 已用 exp 轉過的
fit.link = predict(glm.PO, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "link") # log(mu_i)
res.pear = residuals(glm.PO, type = "pearson")
res.work = residuals(glm.PO, type = "working")

residual_df = data.frame(VILLCODE = Variable_df$VILLCODE,
                         COUNTYNAME = Variable_df$COUNTYNAME,
                         TOWNNAME = Variable_df$TOWNNAME,
                         VILLNAME = Variable_df$VILLNAME,
                         X = Variable_df$X,
                         Y = Variable_df$Y,
                         Nr= Variable_df$Nr,
                         Cluster = Variable_df$Cluster,
                         glm.PO.res.pear =res.pear,
                         glm.PO.fit = fit.resp)
```

### GLM_Poisson 
```{r glm.PO}
glm.qua.PO <- glm(f,  family = quasipoisson(link = "log"), data = Variable_df[c(col_y, col_X_03, col_offset)])
summary(glm.qua.PO)

fit.resp = predict(glm.qua.PO, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "response") # 已用 exp 轉過的
fit.link = predict(glm.qua.PO, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "link") # log(mu_i)
res.pear = residuals(glm.qua.PO, type = "pearson")
res.work = residuals(glm.qua.PO, type = "working")

residual_df["glm.qua.PO.res.pear"] = res.pear
residual_df["glm.qua.PO.fit"] = fit.resp
```
### GLM_NB
- https://bookdown.org/roback/bookdown-BeyondMLR/ch-poissonreg.html
- https://fukamilab.github.io/BIO202/04-C-zero-data.html

先只用各里寺廟、學校、診所密度與男女比、里平均收入建模

```{r glm.NB}
glm.NB<- glm.nb(f ,data = Variable_df[c(col_X_03, col_y, col_offset)])

summary(glm.NB)

fit.resp = predict(glm.NB, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "response") # 已用 exp 轉過的
fit.link = predict(glm.NB, newdata = Variable_df[c(col_y, col_X_03, col_offset)], type = "link") # log(mu_i)
res.pear = residuals(glm.NB, type = "pearson")
res.work = residuals(glm.NB, type = "working")

residual_df["glm.NB.res.pear"] = res.pear
residual_df["glm.NB.fit"] = fit.resp

```

### GWPR

- http://www-sre.wu.ac.at/ersa/ersaconfs/ersa13/ERSA2013_paper_01142.pdf
- 帶寬 (依照論文來定)

```{r}
DM<-gw.dist(dp.locat=data.matrix(((Variable_df[c('X', "Y")]))))

# bw.gwr <- bw.ggwr(f,
#                  data = Variable_shp[c(col_X_03, col_y, "Area_sqkm")],
#                  family = "poisson",
#                  approach = "AICc",
#                  kernel = "gaussian",
#                  adaptive = TRUE,
#                  dMat = DM )
core.HR = mean(.3+.3+.33+.25)*10000; ext.HR = mean(7.7+5.7+5.7+5.6)*10000 
core.bw.gwr = (core.HR/pi)^.5; ext.bw.gwr = (ext.HR/pi)^.5; 
```

```{r}
gwr.PO <- ggwr.basic(f, 
                      data =Variable_shp[c(col_X_03, col_y, "Area_sqkm")],
                      family = "poisson",
                      bw = ext.bw.gwr, 
                      kernel = "gaussian",
                      adaptive = TRUE,
                      dMat = DM)
res.pear = (gwr.PO$SDF$yhat-gwr.PO$SDF$residual)/(gwr.PO$SDF$yhat)^.5
residual_df["gwr.PO.res.pear"] = res.pear
residual_df["gwr.PO.fit"] = gwr.PO$SDF$yhat
summary(gwr.PO$SDF)
```


### Poisson-CAR
- https://carleton.ca/math/wp-content/uploads/Vince-Yordanov-Honors-Project-January-8-2020.pdf
**介紹很多空間模型**
- https://mc-stan.org/users/documentation/case-studies/mbjoseph-CARStan.html

```{r}
```


### Poisson-SAR 
- https://aip.scitation.org/doi/pdf/10.1063/5.0041857
- https://rdrr.io/github/gregmacfarlane/sppois/man/sarpoisson.html

```{r install sppois}
# install.packages("remotes")
# remotes::install_github("gregmacfarlane/sppois")
```

- https://blog.csdn.net/weixin_54000907/article/details/116247097 
**如何在R语言中计算空间自相关指数**
```{r library sppois}
library(sppois)

# 使用距離定義
Variable_sf = st_as_sf(Variable_shp)
ncpoint <- st_centroid(st_geometry(Variable_sf))

nb.k <- knn2nb(knearneigh(ncpoint))
max = max(unlist(nbdists(nb.k, ncpoint)))
nb.dist <- dnearneigh(ncpoint, 0, 0.5*max)
nb.dist2 <- dnearneigh(ncpoint, 0, max)
nb.dist3 <- dnearneigh(ncpoint, 0, 1.5*max)
```

### GLM_poisson with spatial (純粹看不懂) 

`glmmfields`:
- https://github.com/seananderson/glmmfields
- https://cran.r-project.org/web/packages/glmmfields/glmmfields.pdf 
- https://www.mendeley.com/reference-manager/reader/ff1905f3-c0cf-38a2-a752-af1e7632d6eb/2c200ea5-4352-f33f-4851-2b712e71ce53
**Black swans in space: modeling spatiotemporal processes with extremes**

`spaMM`:
- https://kimura.univ-montp2.fr/~rousset/spaMM/spaMMintro.pdf 
- https://www.diva-portal.org/smash/get/diva2:685966/FULLTEXT02
**GLMM 與 GMRF 最大的差別就是後者的隨機效應並不獨立**
- https://www.r-bloggers.com/2019/09/spatial-regression-in-r-part-1-spamm-vs-glmmtmb/
**簡單介紹spaMM**

先只用各里寺廟、學校、診所密度與男女比、里平均收入建模

```{r}
library(glmmfields)
glm.PO.S <- glmmfields(f,
  data = Variable_df[c(col_y, col_X_03, col_offset, "X", "Y")],
  family = poisson(link = "log"),
  lat = "Y", lon = "X", 
  nknots = 50, 
  iter = 500, 
  chains = 1,
  prior_intercept = student_t(3, 0, 10), 
  prior_beta = student_t(3, 0, 3),
  prior_sigma = half_t(3, 0, 3),
  prior_gp_theta = half_t(3, 0, 10),
  prior_gp_sigma = half_t(3, 0, 3),
  seed = 123 # passed to rstan::sampling()
)

summary(glm.PO.S$model@inits)
```

```{r}
p1 = plot(glm.PO.S, type = "prediction", link = F) ; p2 = plot(glm.PO.S, type = "prediction", link = T)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r}
p1 = plot(glm.PO.S, type = "spatial-residual", link = T); p2 = plot(glm.PO.S, type = "spatial-residual", link = F)
ggpubr::ggarrange(p1, p2, ncol = 2)

```

```{r}
plot(glm.PO.S, type = "residual-vs-fitted")
```
### Other models
- file:///C:/Users/ChenWenRu/Downloads/imet131-j-chapitre-6.pdf

## Residual

用 pearson residual

Poisson Model:

$$r^d_i = \frac{\hat{y_i} - \hat{\mu_i}}{\sqrt{\hat{\mu_i}}}$$
Negative Binomial Model:
- https://stats.stackexchange.com/questions/99406/why-are-pearsons-residuals-from-a-negative-binomial-regression-smaller-than-tho

$$r^d_i = ({\hat{y_i} - \bar{\mu_i})}/{\sqrt{\bar{\mu_i}-\frac{\tilde{\mu}}{\theta}}}$$


### Plot

```{r glm.PO.res.pear}
res = residual_df$glm.PO.res.pear
p1 = pdf_plot(res,res<100); p2 = point_plot(res,res<100)
ggpubr::ggarrange(p1, p2, ncol = 2)
mean(res[res<100])
var(res[res<100])
```

```{r glm.qua.PO.res.pear}
res = residual_df$glm.qua.PO.res.pear
p1 = pdf_plot(res); p2 = point_plot(res)
ggpubr::ggarrange(p1, p2, ncol = 2)
mean(res[res<100])
var(res[res<100])
```

```{r glm.NB.res.pear}
res = residual_df$glm.NB.res.pear
p1 = pdf_plot(res,res<5); p2 = point_plot(res,res<5)
ggpubr::ggarrange(p1, p2, ncol = 2)
```

```{r gwr.PO.res}
res = residual_df$gwr.PO.res.pear
p1 = pdf_plot(res); p2 = point_plot(res)
ggpubr::ggarrange(p1, p2, ncol = 2)
```


### Variogram

```{r}
## Transform the dataframe into a SpatialPointDataFrame
## The objective is to obtain a spatial object
coordinates(residual_df)=~X+Y
class(residual_df)
```
```{r glm.PO.res.pear}
idx = residual_df$glm.PO.res.pear<100
Vario_glm.PO = variogram(glm.PO.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_glm.PO)

```

```{r glm.NB.res.pear}
idx = residual_df$glm.NB.res.pear<100
Vario_glm.NB=variogram(glm.NB.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_glm.NB)
```


```{r gwr.PO.res}
idx = residual_df$gwr.PO.res.pear>-90
Vario_gwr.PO=variogram(gwr.PO.res.pear~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_gwr.PO)
```
```{r Nr}
idx = residual_df$gwr.PO.res.pear>-900
Vario_Nr =variogram(Nr~1, ## Here, we assume that there is a constant trend in the data.
                           ## It would not have been the case for an elevation study along a hillslope where 
                           ##there would have been a clear elevation trend regarding the spatial coordinates (along the hillslope)
                     data=residual_df[idx,])
plot(Vario_Nr)
```

### Check the original data

把殘差過大的地方拿出來看看

```{r}
villcode = residual_df$VILLCODE[residual_df$glm.NB.2.res>5]
dt = (Variable_df[Variable_df$VILLCODE %in% villcode ,c('VILLCODE','COUNTYNAME','TOWNNAME','VILLNAME',col_X_03, col_y)])

```